Formulating and Validating RDF Constraints Generically

### 2015-02-11 paper/paper 2/paper 2.tex
4

%\section{Transformations and Automatic Validation of Specific Constraints}
%\section{Transformations between Specific Constraints (Expressed by Different Constraint Languages)}
\section{Transformations between Specific Constraints}
\label{sec:transformations}

RDF data providers ensure that their data conforms to constraints which are part of the contract between the providers and the consumers of that data.
As there is no standard way, these constraints may be represented by multiple constraint languages (there are even more than the five promising ones\footnote{i.a. Bibframe, DQTP, Pellet ICV, RDFUnit, SPARQL, Stardog ICV}).
There is an obvious need to transform a specific constraint (\ms{$sc_{\alpha}$}) (expressed by any constraint language \ms{$\alpha$}) into a specific constraint (\ms{$sc_{\beta}$}) (expressed by any constraint language \ms{$\beta$}) - when they have the same semantics.
We use the RDF-CO as intermediate representation for these transformations.
If we define mappings ($m(sc_{\alpha}), m(sc_{\beta})$) from two equivalent specific constraints to the corresponding generic constraint (\ms{gc}),
we can convert them automatically:

@@

We do not need to define mappings for each pair of semantically equivalent specific constraints.
Let's assume we are able to express a constraint by 5/6/7/8/9/10 constraint languages.
Without the intermediate generic constraint, we would have to define one mapping for each pair of specific constraints expressed by a different constraint languages
- that are \ms{\( {n \choose 2} \)} mappings for each constraint (10/15/21/28/36/45 mappings).
With an intermediate generic constraint, we only need to define \ms{n} mappings (one mapping for each specific constraint) from the specific constraints to the generic constraint (5/6/7/8/9/10 mappings).

@@

\ms{Allowed Values}\footnote{corresponds to \emph{R-30-ALLOWED-VALUES-FOR-RDF-OBJECTS} and \emph{R-37-ALLOWED-VALUES-FOR-RDF-LITERALS}}
is a common requirement to narrow down the value space of a property by an exhaustive enumeration of the valid values (both literals or resources). 
The Enterprise, e.g., is only commanded by the captains Archer, Kirk, and Picard.
This complex constraint is composed of the two simple constraints \ms{universal quantification} and \ms{allowed values} which can be represented in DL:

@@

\begin{DL}
Enterprise $\equiv$ $\forall$ commandedBy . \{Archer\} $\sqcup$ \{Kirk\} $\sqcup$ \{Picard\}
\end{DL}

First, we map these DL statements to two simple generic constraints corresponding to the RDF-CO (we define the class \ms{EnterpriseCaptains} by simply enumerating its instances):

@@

\begin{gcotable}
property & Enterprise & commandedBy & - & EnterpriseCaptains & $\forall$ & - \\
class & EnterpriseCaptains & - & - & \{Archer\},\{Kirk\},\{Picard\} & $\sqcup$ & - \\
\end{gcotable}

The next step is to define one mapping for each specific constraint to the generic constraint.
The specific constraints (\ms{$sc_{OWL 2}$}, \ms{$sc_{ReSh}$}, and \ms{$sc_{ShEx}$}) expressed by OWL 2, ReSh, and ShEx look like this:

\begin{ex}
# OWL 2:
Enterprise rdfs:subClassOf [ a owl:Restriction .
    owl:onProperty commandedBy .
    owl:allValuesFrom [ a owl:Class ;
        owl:oneOf ( Archer Kirk Picard ) ] ] .
# ReSh:
Enterprise a rs:ResourceShape ; rs:property [
    rs:name "commandedBy" ; rs:propertyDefinition commandedBy ;
    rs:allowedValue Archer , Kirk , Picard ;
    rs:occurs rs:Exactly-one ; ] .
# ShEx:
Enterprise {
    commandedBy (Archer Kirk Picard) }
\end{ex}

@@

\section{Validation of Specific and Generic Constraints}
\label{sec:validation}

>>>>

%\section{Transformations and Automatic Validation of Specific Constraints}
%\section{Transformations between Specific Constraints (Expressed by Different Constraint Languages)}
%\section{Transformations between Specific Constraints}
\section{How to Improve the Interoperability of Constraints}
\label{sec:transformations}

RDF data providers have to ensure that their data conforms to constraints which are part of explicitly or implicitly concluded contracts between data providers and consumers.
As there is no standard way, these constraints may be represented by a variety of possible constraint languages (there are even more than the five promising ones\footnote{i.a. Bibframe, Data Quality Test Patterns, Pellet ICV, RDFUnit, SPARQL, Stardog ICV}).
There is an obvious need to transform any \emph{specific constraint} (\emph{$sc_{\alpha}$}) (expressed by any constraint language \emph{$\alpha$}) into any other \emph{specific constraint} (\emph{$sc_{\beta}$}) (expressed by any constraint language \emph{$\beta$}) - when they have the same semantics.
We use the \emph{RDF-CV} as an intermediate representation vocabulary for these transformations.
If we define mappings ($m(sc_{\alpha}), m(sc_{\beta})$) between equivalent \emph{specific constraints} and the corresponding \emph{generic constraint} (\emph{gc}),
we can convert them automatically:

@@

We do not need to define mappings for each pair of semantically equivalent \emph{specific constraints}.
Let's assume we are able to express a constraint type by 5/6/7/8/9/10 constraint languages.
Without the intermediate \emph{generic constraint}, we would have to define for each constraint type \emph{1} mapping for each pair of \emph{specific constraints} (expressed by different constraint languages)
- that are \emph{\( {n \choose 2} \)} mappings (10/15/21/28/36/45 mappings).
With an intermediate \emph{generic constraint}, however, we only need to define for each constraint type \emph{n} mappings (one mapping for each \emph{specific constraint}) from \emph{n} \emph{specific constraints} to the corresponding \emph{generic constraint} (5/6/7/8/9/10 mappings).

@@

The {\em existential quantification}\footnote{Corresponds to {\em R-86-EXISTENTIAL-QUANTIFICATION-ON-PROPERTIES}}
\ms{Publication $\equiv$ $\exists$ publisher.Publisher} restricts publications to have at least one publisher and can be expressed by at least 10 constraint languages\footnote{at least by OWL2, SPARQL, ShEx, Bibframe, DSP, ReSh, SPIN, DQTP, Pellet IVC, Stardog ICV}:

\begin{ex}
# OWL 2:
[ a owl:Restriction ;
  owl:onProperty publisher ;
  owl:someValuesFrom Publisher ;
  rdfs:subClassOf Publication ] .

# SPARQL:
ASK { ?x a Publication . FILTER NOT EXISTS { ?x publisher [ a Publisher ] } . }

# ShEx:
Publication {
    publisher @Publisher{1,} }

# DSP:
descriptionTemplate a dsp:DescriptionTemplate ;
    dsp:resourceClass Publication ; 
    dsp:statementTemplate [ a dsp:NonLiteralStatementTemplate ;
        dsp:minOccur 1 ; dsp:maxOccur "infinity" ; 
        dsp:property publisher ; 
        dsp:nonLiteralConstraint [ a dsp:NonLiteralConstraint ;
            dsp:valueClass Publisher ] ] .

# ReSh:
Publication a rs:ResourceShape ; rs:property [
    rs:propertyDefinition :publisher ;
    rs:valueShape :Publisher ;
    se:min 1 ; se:maxundefined true ; ] .

# SPIN:
CONSTRUCT {
    _:cv a spin:ConstraintViolation ; spin:violationRoot ?subject ;
         rdfs:label ?violationMessage ; spin:violationPath ?property . } WHERE 
{ ?subject a ?classSubject . FILTER NOT EXISTS { ?subject ?property [ a ?classObject ] } . }
\end{ex}

%\emph{Allowed Values}\footnote{corresponds to \emph{R-30-ALLOWED-VALUES-FOR-RDF-OBJECTS} and \emph{R-37-ALLOWED-VALUES-FOR-RDF-LITERALS}}
%is a common requirement (constraint type) to narrow down the value space of a property by an exhaustive enumeration of the valid values (both literals or resources). 
%The Enterprise, e.g., is only commanded by the captains Archer, Kirk, and Picard.
%This \emph{complex constraint} is composed of the two \emph{simple constraints} \emph{universal quantification} and \emph{allowed values} which can be represented in DL:

@@

%\begin{DL}
%Enterprise $\equiv$ $\forall$ commandedBy . \{Archer\} $\sqcup$ \{Kirk\} $\sqcup$ \{Picard\}
%\end{DL}

%First, we map these DL statements to two \emph{simple generic constraints} corresponding to the \emph{RDF-CV} (we define the class \ms{EnterpriseCaptains} by simply enumerating its instances):

First, the DL statement \ms{Publication $\equiv$ $\exists$ publisher.Publisher} is mapped to a \emph{simple generic constraint} according to the \emph{RDF-CV} (see table \ref{tab:existential-quantification-as-simple-constraints}).

@@

%\begin{gcotable}
%property & Enterprise & commandedBy & - & EnterpriseCaptains & $\forall$ & - \\
%class & EnterpriseCaptains & - & - & \{Archer\},\{Kirk\},\{Picard\} & $\sqcup$ & - \\
%\end{gcotable}

\begin{table}
  \scriptsize
  \sffamily
  \vspace{0cm}
	\centering
		\begin{tabular}{l|l|l|l|l|l|l}
      \textbf{c. type} & \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
property & Publication & publisher & - & Publisher & $\exists$ \\
		\end{tabular}
	\caption{Existential Quantification as Simple Constraints}
	\label{tab:existential-quantification-as-simple-constraints}
\end{table}

The next step is to define one mapping for each \emph{specific constraint} (\emph{$sc_{OWL 2}$}, \emph{$sc_{ReSh}$}, \emph{$sc_{ShEx}$}, ...) to the \emph{generic constraint}.

%\begin{ex}
%# OWL 2:
%Enterprise rdfs:subClassOf [ a owl:Restriction .
    %owl:onProperty commandedBy .
    %owl:allValuesFrom [ a owl:Class ;
        %owl:oneOf ( Archer Kirk Picard ) ] ] .
%# ReSh:
%Enterprise a rs:ResourceShape ; rs:property [
    %rs:name "commandedBy" ; rs:propertyDefinition commandedBy ;
    %rs:allowedValue Archer , Kirk , Picard ;
    %rs:occurs rs:Exactly-one ; ] .
%# ShEx:
%Enterprise {
    %commandedBy (Archer Kirk Picard) }
%\end{ex}

@@

%\section{Validation of Specific and Generic Constraints}
\section{How to Provide Constraint Validation Out-Of-The-Box}
\label{sec:validation}
